---
title: "Homework 3"
author: E. Brennan Bollman
date: '`r format(Sys.time(), "%y-%m-%d")`'
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom")) 

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d()
scale_fill_discrete = scale_fill_viridis_d()
```

## Problem 1

##### Step 1: describe the instacart dataset. 

```{r}
data("instacart")
instacart
```

This dataset contains `r nrow(instacart)` rows of individual product orders from Instacart and `r ncol(instacart)` columns of variables such as: user ID, product ID, day and hour of the week at which the order was placed, as well as product names and the departments / aisles in which they can be found. The dataset contains information from `r instacart %>% distinct(user_id) %>% count()` unique users who ordered `r instacart %>% distinct(product_id) %>% count()` products. 


Number of aisles and which have most items ordered from.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered, arranging sensibly.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table. 
Step 1  of this is to filter these aisles and count most popular items in these aisles. Step 2: rank within each of the 3 aisles.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table). Step 1: look only at those items. Step 2: summarize to find mean hour of day at which each item ordered. Step 3: find that mean for each day of week

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```



## Problem 2

## Problem 3